---
title: "Using seaice"
author: "Michael D. Sumner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using seaice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The seaice package aims to provide easy access to the most comprehensive data sets of sea ice concentration data available via passive microwave remote sensing. 

Please see the vignette for obtaining the data set/s, `r vignette("get_seaice", package = "seaice")`. 

An easy first step is to read the latest data. 

```{r, eval = FALSE}
library(seaice)
southern <- read_south_seaice()
northern <- read_north_seaice()

op <- par(mfrow = c(2, 1))
plot(southern, col = palr::icePal(100))
plot(northern, col = palr::icePal(100))
par(op)

southern 
northern
```

The object returned is a `RasterLayer` from the `raster` package. Note that each object is specific to its hemisphere, these are different map projections and they cannot be plotted together natively. 

The read functions have an argument `latest` set to `TRUE` by default, and so the date of data returned will depend on the last synchronization run of the data, as described in the `get_seaice` data vignette. 
